### 2023-05-19 paper/localpsf.tex
some writing updates

Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following approximations:
\begin{align}
\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \label{eq:translate3}\\
\phi_x(y)/\spatialvol(x) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) / \spatialvol(x'). \label{eq:translate4}
\end{align}
These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones. Our method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations. Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close. Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation~\cite{MahoneyDrineas09}. However, the required rank in the low rank approximation can be large, which makes algorithms based on~\eqref{eq:translate1} expensive. Operators that satisfy~\eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$. It is straightforward to show that if equality holds in~\eqref{eq:translate2}, then $\Aop$ is a convolution operator. Locally translation invariant operators act like convolutions locally, and can therefore be well approximated by PC approximations.

@@

We use approximation~\eqref{eq:translate4}, which is the same as~\eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side. These factors make the approximation more robust if $\spatialvol(x)$ varies widely. Approximation~\eqref{eq:translate4} is equivalent to~\eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$. We call~\eqref{eq:translate4} ``normalized local mean displacement invariance.''

>>>>

Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following approximations:
\begin{align}
\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \label{eq:translate3}\\
\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \spatialvol(x) / \spatialvol(x'). \label{eq:translate4}
\end{align}
These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones. Our method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations. Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close. Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation~\cite{MahoneyDrineas09}. However, the required rank in the low rank approximation can be large, which makes algorithms based on~\eqref{eq:translate1} expensive. Operators that satisfy~\eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$. It is straightforward to show that if equality holds in~\eqref{eq:translate2}, then $\Aop$ is a convolution operator. Locally translation invariant operators act like convolutions locally, and can therefore be well approximated by PC approximations.

@@

We use approximation~\eqref{eq:translate4}, which is the same as~\eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side. These factors make the approximation more robust if $\spatialvol(x)$ varies widely. Approximation~\eqref{eq:translate4} is equivalent to~\eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$. We call~\eqref{eq:translate4} \emph{normalized local mean displacement invariance}.

We have experimented with more advanced interpolation schemes involving $\Sigma$, in which one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point. We do not use these more advanced schemes because we find that they are much less robust to negative entries in the integral kernel. Although we assume the integral kernel has non-negative entries (Condition 3 in Section \ref{sec:conditions_2}), in practice the integral kernel typically some negative entries due to discretization error.

%but we find they are not robust, and therefore we do not use them:
%\begin{align*}
%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \spatialvol(x) / \spatialvol(x') \\
%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \left( \spatialvol(x) \det{\Sigma^{1/2}(x')} \right) / \left( \spatialvol(x') \det{\Sigma^{1/2}(x)} \right)
%\end{align*}
%In these schemes, one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point. We find that computation of $\Sigma$ is more susceptible to numerical error than computation of $V$ and $\mu$, in the sense that spurious negative numbers in $\phi_x$ cause $\Sigma(x)$ to become highly ill-conditioned or even indefinite. When $\Sigma$ is used for picking sample points (Section \ref{sec:sample_point_selection}), we simply avoid picking sample points $p$ with ``bad'' $\Sigma(p)$. These schemes require using $\Sigma_x$ at arbitrary $x$, rather than just the sample points. When picking sample points, one may
